\documentclass[journal]{IEEEtran}

% *** PACKAGES ***
% newtxtext gives Times with full shape coverage (incl. TS1/sc
% and OT1/scit that stock IEEEtran Times omits, silencing the
% font-shape-undefined warnings).
\usepackage{newtxtext}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{url}
% hyperref must come AFTER IEEEtran's own packages but BEFORE
% any custom defs; hidelinks suppresses the default colored
% boxes around links. Fixes \url{} truncation when URLs break
% across lines (links now stay glued together end-to-end).
\usepackage[hidelinks,breaklinks=true]{hyperref}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta,shapes.geometric,calc,patterns}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepgfplotslibrary{statistics}

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

% ---------------------------------------------------------------
% TITLE
% ---------------------------------------------------------------
\title{LamQuant Lossless: State-of-the-Art Lossless EEG Compression
Designed for Microcontroller Deployment}

% ---------------------------------------------------------------
% AUTHOR
% ---------------------------------------------------------------
\author{Brian~Lam,~\IEEEmembership{Student Member,~IEEE}%
\thanks{B. Lam is the founder of OpenHuman Technologies LLC,
Florida, USA, and a Student Member of the IEEE Florida West
Coast Section, Region 3 (member \#102403636). E-mail:
briankhanglam@usf.edu. ORCID: 0009-0001-5463-2324.
Patent Pending US \#64/032,641.
Manuscript received May 26, 2026.}}

\markboth{IEEE Transactions on Biomedical Circuits and Systems}%
{Lam: LamQuant Lossless EEG Compression}

\maketitle

% ---------------------------------------------------------------
% ABSTRACT
% ---------------------------------------------------------------
\begin{abstract}
Lossless compression of clinical electroencephalography (EEG) is a
prerequisite for wireless transmission and long-term archival, yet
existing approaches either fail to exploit EEG signal structure or
are not deployable on microcontroller hardware. We present LamQuant
Lossless, an integer-only EEG codec combining a Le~Gall 5/3 lifting
discrete wavelet transform, per-subband adaptive linear prediction
(orders 3, 3, 6, 8 with an optional max-order-16 mode), bias
cancellation, and Golomb-Rice entropy coding, in a wire format
that runs byte-identically on x86, ARM, and RISC-V without
floating-point hardware. On the CHB-MIT Scalp EEG Database under
parameters identical to Chen et~al.~\cite{chen2018vlsi}, LamQuant
Lossless achieves 2.72:1---a 15.9\,\% improvement over the prior
state of the art. On the Temple University EEG Corpus v2.0.2
(1.76\,TB, 70{,}831 EDF files per the corpus
AAREADME~\cite{tueg_aareadme_v202}), it achieves 2.287:1,
establishing the first, to our knowledge, lossless compression
benchmark on the largest open clinical EEG corpus by total bytes
in existence. Bit-exact reconstruction was verified across
88{,}147 encode/decode operations across thirteen clinical and
research EEG datasets with zero failures.
\end{abstract}

% ---------------------------------------------------------------
% KEYWORDS
% ---------------------------------------------------------------
\begin{IEEEkeywords}
EEG compression, lossless coding, Golomb-Rice entropy coding,
integer lifting wavelet transform, linear predictive coding,
microcontroller deployment, Temple University EEG Corpus,
wireless EEG, clinical archival
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

% ---------------------------------------------------------------
% I. INTRODUCTION
% ---------------------------------------------------------------
\section{Introduction}
\label{sec:introduction}

\IEEEPARstart{W}{ireless} EEG acquisition imposes a bandwidth
constraint that grows quickly with channel count, sample rate,
and bit depth. A 32-channel ambulatory system at 1024\,Hz
(the upper end of the TUEG sample-rate range~\cite{tueg_aareadme_v201})
with 24-bit resolution generates $32 \times 1024 \times 24 =
786$\,kbps. A 256-channel high-density research system at
1\,kSPS with 24-bit resolution generates $256 \times 1024 \times
24 \approx 6.3$\,Mbps, and a clinical intracranial system at
16\,kSPS with 24-bit resolution over 32 channels generates
$32 \times 16000 \times 24 \approx 12$\,Mbps. Bluetooth Low Energy~5.x sustains
$\approx$1.4\,Mbps application-layer; high-density and intracranial
configurations exceed it, and even modest ambulatory loads leave
little headroom for retransmission, encryption, and concurrent
telemetry uncompressed.
Lossless compression therefore enables the wider range of
clinical configurations to fit a low-power wireless transport,
and the bit-exact constraint is hard: medicolegal and FDA archival
requirements forbid silent sample modification. For archival and medicolegal purposes the compression must
satisfy PRD\,=\,0\%: every sample must be recovered bit-exactly, and
every annotation, timestamp, and sidecar file must be preserved with
forensic fidelity.

General-purpose compressors such as gzip and zstd exploit
byte-level string repetition and ignore the statistical structure
of EEG signals. On TUEG v2.0.1, gzip~$-9$ achieves 1.6033:1
(measured directly on the corpus production server via
\texttt{tools/bench\_gzip\_baseline.sh}, 6.02\,GB $\rightarrow$
3.75\,GB over 456 EDFs of TUEG v2.0.2 edf/000, 274\,s wall).
Signal-aware approaches based on predictive coding and entropy
coding have been studied since the 1990s~\cite{memon1998}. The
current state of the art for hardware-oriented lossless EEG
compression is Chen et~al.~\cite{chen2018vlsi} at 2.35:1 on the
CHB-MIT Scalp EEG Database, using a fixed fuzzy predictor and
tri-stage Huffman/Golomb-Rice encoder implemented in
0.18\,$\mu$m CMOS. No prior work has simultaneously achieved
state-of-the-art compression ratio, microcontroller deployability
on commodity RISC-V hardware, and validation at the scale of a
full clinical corpus.

This paper presents LamQuant Lossless, the archival compression layer
of the OH-EEG open-source wireless EEG platform. The codec applies a
Le~Gall 5/3 integer lifting wavelet transform~\cite{sweldens1996,itu_t800},
linear-predictive coding with a fixed per-subband order schedule of
$\{3,3,6,8\}$ (and an optional adaptive mode of maximum order 16)
combined with bias cancellation~\cite{sriraam2008}, and Golomb-Rice
entropy coding in a fixed integer-only wire format deployable on any
x86, ARM, or RISC-V target without floating-point hardware. Evaluated on CHB-MIT under identical parameters to Chen et~al.,
LamQuant Lossless achieves 2.72:1 CR---a 15.9\,\% improvement
over the prior state of the art. On TUEG v2.0.2 (released
2026-05-21; 1.76\,TB, 70{,}831 EDF files per the corpus
AAREADME~\cite{tueg_aareadme_v202}), it achieves 2.287:1. This
establishes a lossless compression benchmark on the largest open
clinical EEG dataset in existence. Bit-exact reconstruction was
verified across 88{,}147 encode/decode operations on 74{,}422
unique recordings spanning thirteen clinical and research EEG
datasets (the seven TUH corpora, CHB-MIT, PhysioNet Siena Scalp
EEG, Sleep-EDF, CAP Sleep, EEG Motor Movement/Imagery, and
OpenNeuro ds004100), with zero failures.

The remainder of this paper is organized as follows:
Section~\ref{sec:background} reviews prior work and establishes
theoretical bounds; Section~\ref{sec:algorithm} describes the
LamQuant Lossless algorithm and wire format;
Section~\ref{sec:results} presents experimental results;
Section~\ref{sec:discussion} discusses implications for clinical
deployment; and Section~\ref{sec:conclusion} concludes.


% ---------------------------------------------------------------
% II. BACKGROUND
% ---------------------------------------------------------------
\section{Background}
\label{sec:background}

\subsection{Clinical EEG Storage and the EDF Format}
The European Data Format (EDF) and its extensions EDF+ and BDF are
the dominant standards for clinical EEG storage~\cite{kemp1992}. An
EDF file encodes fixed-duration data records containing multiplexed
integer samples from each channel alongside a structured ASCII header
and, in EDF+, embedded annotations. The integer sample representation
and channel-multiplexed layout create a signal structure that
general-purpose compressors cannot efficiently exploit.

\subsection{The Temple University EEG Corpus}
The Temple University EEG Corpus (TUEG) is the largest open clinical
EEG repository by total bytes~\cite{obeid2016}. Per the corpus
AAREADME~\cite{tueg_aareadme_v202}, v2.0.2 (released 2026-05-21)
contains 14{,}987 patients, 26{,}846 sessions, 70{,}831 EDF files
(27{,}074.4 hours). Sample frequencies range from 250\,Hz to 1024\,Hz
and approximately 95\% of recordings include a 10/20 montage
subset. Direct measurement on the local mirror sums to
1{,}756{,}355{,}590{,}458\,bytes (1.76\,TB), spanning
diverse montages, sampling rates, and recording conditions
representative of a large academic epilepsy monitoring unit (EMU).
Two files present in the earlier v2.0.0 release were identified as
corrupted and rectified in v2.0.1; both were processed and flagged
correctly by LamQuant Lossless. The scale and clinical diversity of
TUEG make it a substantially harder compression target than curated
laboratory datasets, and no prior lossless compression benchmark on
the full corpus has been published.

\subsection{Limitations of General-Purpose Compression on EEG}
Gzip (DEFLATE: LZ77 + Huffman)~\cite{ziv1977,deutsch1996} and
zstd target byte-level repetition and dictionary-based string
matching. EEG signals produced by
delta-sigma ADCs exhibit inter-sample temporal correlation and
inter-channel spatial correlation that these compressors do not
model. Applied to TUEG edf/000, gzip $-9$ achieves 1.6033:1 CR,
confirmed by direct measurement (456 files, 6.02\,GB raw, 3.75\,GB
compressed, 274\,s wall on the corpus production server). LamQuant
Lossless achieves 2.299:1 on the same 456-file subset (measured
by \texttt{tools/bench\_per\_file\_cr.py})---43\,\% above the gzip
baseline---by exploiting signal
structure through predictive coding and signal-matched entropy coding.

A more competitive comparator is FLAC (version 1.5.0 used in this
work; the most recent stable release at submission), the de-facto
lossless audio codec, whose architecture (per-frame fixed or
windowed LPC up to order 12 at \texttt{-7}, order 32 at
\texttt{-8}, followed by partitioned Rice coding) is structurally
similar to LamQuant's.
A fair per-channel FLAC bench (each EDF channel encoded as a
separate \texttt{flac --bps=16 --channels=1} stream so FLAC's
stereo-decorrelation tooling does not penalise the multi-channel
input) yields the numbers shown in Table~\ref{tab:comparison}: on
TUEG edf/000, FLAC \texttt{-7} achieves 2.2574:1 and FLAC
\texttt{-8} 2.2609:1, against LamQuant's 2.299:1 ($+$1.7\,\% over
FLAC \texttt{-8}); on CHB-MIT, FLAC \texttt{-7} achieves 2.6926:1
and FLAC \texttt{-8} 2.6945:1 against LamQuant's 2.7229:1
($+$1.1\,\% over FLAC \texttt{-8}). FLAC \texttt{-7} $\rightarrow$
\texttt{-8} (LPC ceiling $12 \rightarrow 32$) gains less than
0.2\,\% on either corpus, confirming that EEG autocorrelation
saturates LPC at order $\sim$8--12 and that raising the
predictor ceiling further is not the lever; this is consistent
with an internal P$_{\max}$ ablation showing byte-identical
output at LamQuant P$_{\max} \in \{16, 32, 128\}$. The
1--2\,\% raw-CR edge LamQuant retains over FLAC \texttt{-8}
comes from the four-band lifting DWT (FLAC has no transform stage)
and per-subband AIC order selection, but the headline contribution
relative to FLAC is not the CR delta: it is microcontroller
deployability. FLAC's reference encoder does not target sub-MB
SRAM real-time operation on a 150~MHz RISC-V core; LamQuant does,
and the wire format is byte-identical across x86, ARM, and RISC-V
backends (Section~\ref{sec:results}).

\subsection{Shannon Entropy Bound for Clinical EEG}
The compression ceiling for a lossless codec is set by the Shannon
entropy of the residuals it actually encodes. Empirical entropies
measured on a 200-file random sample of TUEG~v2.0.1
(\texttt{tools/bench\_shannon\_entropy.py}, 2.45\,$\times$\,10$^9$
samples) are:

\begin{itemize}
\item \textbf{$H_0(X)$ = 10.775\,bits/sample} on raw 16-bit
  digital samples~$\Rightarrow$ a memory-less entropy coder cannot
  exceed $16/10.775 \approx 1.485{:}1$ CR.
\item \textbf{$H_0(\Delta X)$ = 9.337\,bits/sample} on the
  first-order difference $X[t]-X[t-1]$ (a fixed-coefficient LPC-1
  predictor)~$\Rightarrow$ a memory-less entropy coder over LPC-1
  residuals cannot exceed $16/9.337 \approx 1.714{:}1$ CR.
\end{itemize}

Both ceilings sit well below the 2.287:1 measured for LamQuant
Lossless on the same corpus. To be precise about what this
comparison means: the $H_0(X)$ and $H_0(\Delta X)$ ceilings are
the achievable compression of a memory-less entropy coder over
the raw alphabet and the LPC-1 residual respectively; LamQuant is
a higher-order predictor (four-band Le~Gall 5/3 lifting DWT
followed by per-subband adaptive LPC of orders 3, 3, 6, 8, with
an optional order-16 mode) and the comparison is therefore
\emph{lower-order ceiling vs higher-order method} by design — not
"ceiling vs achievement" in the same probabilistic class. The
2.287:1 result implies an effective residual entropy of $\leq
16/2.287 \approx 7.00$\,bits/sample after the four-band lifting
DWT plus per-subband LPC pipeline, demonstrating that the
higher-order structure exploited by this pipeline is materially
richer than what LPC-1 captures (a gap of $\geq 9.337 - 7.00 =
2.34$\,bits/sample) and that there is concrete residual entropy
left on the table for future fixed-codebook contextual modelling
beyond Golomb-Rice.

By contrast, lossless compression ratios reported in excess of
$\sim$3:1 on comparable EEG signal classes
\cite{sriraam2012} typically arise from evaluation on
curated, low-noise, single-session laboratory datasets, where
$H_0(X)$ and $H_0(\Delta X)$ are substantially lower than on a
clinically diverse corpus. The 2.287:1 result on TUEG reflects
honest benchmarking on a clinically realistic mixture and beats
both first-order Shannon ceilings on that mixture.

\subsection{Prior Work in Lossless EEG Compression}
Early signal-aware approaches applied predictive coding with
dictionary or context-based entropy coders~\cite{memon1998}.
Subsequent work explored integer wavelet transforms combined with
neural network predictors and Lempel-Ziv arithmetic
encoding~\cite{sriraam2012}, achieving CR\,=\,2.99:1 on a small
single-session dataset. Sriraam and Eswaran~\cite{sriraam2008}
proposed context-based bias cancellation as a lightweight alternative
to neural predictors, demonstrating competitive performance on
single-channel recordings.
Chen et~al.~\cite{chen2018vlsi} advanced the state of the art for
hardware-oriented compression with a VLSI implementation achieving
2.35:1 on CHB-MIT using a fixed fuzzy predictor and tri-stage
Huffman/Golomb-Rice encoder. In the adjacent ECG-on-implantable
space, Higgins et~al.~\cite{higgs2024ecg} reported a
region-adaptive predictor with Golomb-Rice coding tailored to a
low-power region; an LPC + Golomb-Rice composition is therefore
an idiom shared across the biosignal-on-MCU community, and the
LamQuant kernel choice is consistent with current TBME practice
rather than novel. Recent low-power-wireless work by
Cuevas-L{\'o}pez et~al.~\cite{cuevas2022wireless} achieves
$\sim$\,1.5:1 lossless compression of in-vivo neural
electrophysiology on FPGA (delta + Huffman, $\leq$\,3\,mW).
At the bandwidth extreme, He et~al.~\cite{he2024nct} reported a
TBioCAS event-based neural compressive telemetry IC delivering
$>$\,11.4{$\times$} lossless reduction on high-bandwidth
intracortical BCI data at 1\,$\mu$W/channel by exploiting
inter-channel spike-event sparsity. That regime is structurally
distinct from scalp clinical EEG, where adjacent samples are
broadband-correlated rather than sparse-event. None of the cited prior-work systems
validate at the scale of a full clinical corpus: typical evaluations
use single-session datasets ranging from a few hundred megabytes to
tens of gigabytes. Hardware-oriented systems published in the IEEE
TBioCAS / JSSC line~\cite{chen2018vlsi} report compression numbers
measured on the same restricted CHB-MIT subset and do not address
metadata or annotation fidelity, which have become first-order concerns
for clinical archival use and ML applications. LamQuant Lossless builds on this lineage,
replacing the fixed predictor with adaptive LPC (per-subband order
schedule plus optional max-order-16 Levinson--Durbin walk), adopting
the Le~Gall 5/3 lifting transform from JPEG~2000~\cite{sweldens1996,itu_t800},
adding the LMLFOOT1 window offset table for random-access decoding,
and validating at full TUEG corpus scale with byte-exact
verification of every annotation sidecar.


% ---------------------------------------------------------------
% III. ALGORITHM
% ---------------------------------------------------------------
\section{LamQuant Lossless Algorithm}
\label{sec:algorithm}

The LamQuant Lossless encoding pipeline proceeds in four stages
(Fig.~\ref{fig:pipeline}): (1)~integer lifting DWT,
(2)~LPC prediction with bias cancellation, (3)~Golomb-Rice entropy
coding, and (4)~packetization into the LML wire format. All
arithmetic is integer-only; no floating-point operations are
required at any stage.

\begin{figure}[!t]
\centering
\begin{tikzpicture}[
    node distance=4mm and 6mm,
    box/.style={draw, rectangle, rounded corners=2pt,
                minimum height=8mm, minimum width=14mm,
                align=center, font=\scriptsize},
    sig/.style={draw, rectangle, dashed, rounded corners=2pt,
                minimum height=6mm, minimum width=14mm,
                align=center, font=\scriptsize\itshape},
    arr/.style={-{Stealth}, thick},
]
\node[sig] (in)  {EDF$_{\text{in}}$\\21 ch $\times$ $T$ i16};
\node[box, right=of in]   (dwt) {lifting\\5/3 DWT};
\node[box, right=of dwt]  (lpc) {LPC\\predict};
\node[box, below=of lpc]  (bias){bias\\cancel};
\node[box, left=of bias]  (gr)  {Golomb-\\Rice};
\node[box, left=of gr]    (lml) {LML\\packet};
\node[sig, left=of lml]   (out) {.lml\\out};

\draw[arr] (in)  -- (dwt);
\draw[arr] (dwt) -- (lpc);
\draw[arr] (lpc) -- (bias);
\draw[arr] (bias) -- (gr);
\draw[arr] (gr)  -- (lml);
\draw[arr] (lml) -- (out);

% Side annotations
\node[font=\tiny, above=0.5mm of dwt] {Stage~1};
\node[font=\tiny, above=0.5mm of lpc] {Stage~2a};
\node[font=\tiny, below=0.5mm of bias]{Stage~2b};
\node[font=\tiny, below=0.5mm of gr]  {Stage~3};
\node[font=\tiny, below=0.5mm of lml] {Stage~4};
\end{tikzpicture}
\caption{LamQuant Lossless encoding pipeline. All stages operate
on integer samples; the three-level Le~Gall 5/3 lifting DWT,
per-subband LPC with bias cancellation, and Golomb-Rice coder run
sequentially on each 10-second window. Per-window metadata (LPC
coefficients + Rice parameter $k$) is packed into the LML header
alongside the GR payload. The dashed boxes mark the file-level
boundaries; everything between them is integer-only arithmetic
suitable for floating-point-free MCU deployment. (The OH-EEG
firmware optionally applies a biquad high-pass filter as a
\emph{pre-processing} stage before lossless encoding; that filter
is lossy by nature and \emph{not part of the LamQuant Lossless
codec}---the lossless guarantee in this paper applies to the
integer signal presented to the lifting DWT input.)}
\label{fig:pipeline}
\end{figure}

\subsection{Integer Le~Gall 5/3 Lifting DWT}
The Le~Gall 5/3 wavelet filter, adopted in JPEG~2000 for its
lossless integer-to-integer lifting
construction~\cite{sweldens1996,itu_t800}, is applied per channel per window.
The lifting scheme computes predict and update steps using only
integer addition and right-shifts, mapping integer input samples to
integer subband coefficients without rounding error. The decision to
employ an integer lifting approach was motivated by the hardware
constraints of MCU deployment: the target platform (RP2350, Hazard3
RISC-V) does not include a hardware floating-point unit, and
floating-point emulation in software incurs unacceptable latency at
32-channel real-time encoding rates. In-place computation requires no
auxiliary buffer beyond the working window, a critical property for
MCU deployments with constrained SRAM. The transform decorrelates
inter-sample temporal redundancy, concentrating signal energy in the
low-frequency subband and producing sparse high-frequency residuals
that compress efficiently under subsequent entropy coding. The 5/3
filter was selected over longer biorthogonal filters (e.g., 9/7) on
the basis that its shorter support minimizes per-window boundary
overhead while achieving sufficient decorrelation for EEG's
predominantly low-frequency energy distribution.

\noindent\textit{DWT subbands vs.\ EEG clinical bands.}
The DWT is a dyadic frequency split via the analysis filters,
not the standard EEG clinical bands (delta/theta/alpha/beta/gamma).
At 250\,Hz, the three DWT levels produce four subbands: $L_3$
(0--15.6\,Hz; delta + theta + alpha + low beta), $L_2$
(15.6--31.25\,Hz; high beta + low gamma), $L_1$
(31.25--62.5\,Hz; gamma + line-noise harmonics), and $L_0$
(62.5--125\,Hz; EMG / high-frequency artifact). The codec
compresses subbands, not clinical bands; the mapping above is
for orientation only.

\subsection{LPC Prediction with Bias Cancellation}
Linear predictive coding is applied to each DWT subband per channel
per window. The predictor of order $P$ computes:
%
\begin{equation}
e[n] = x[n] - \sum_{i=1}^{P} a_i \, x[n-i]
\label{eq:lpc}
\end{equation}
%
where $x[n]$ are the subband coefficients, $a_i$ are the predictor
coefficients, and $e[n]$ are the prediction residuals stored in the
compressed stream. The codec uses a fixed per-subband order schedule
of $P \in \{3, 3, 6, 8\}$ for the four DWT subbands in the default
Fixed mode, indexed from the highest-frequency detail subband to
the lowest-frequency approximation:
\begin{center}
\begin{tabular}{lcc}
\toprule
DWT subband & Freq.\ at 250\,Hz & LPC order $P$ \\
\midrule
$L_0$ (highest detail)  & 62.5--125\,Hz   & 3 \\
$L_1$                   & 31.25--62.5\,Hz & 3 \\
$L_2$                   & 15.6--31.25\,Hz & 6 \\
$L_3$ (approximation)   & 0--15.6\,Hz     & 8 \\
\bottomrule
\end{tabular}
\end{center}
This schedule reflects the spectral structure of clinical EEG:
the lowest-frequency approximation subband carries the bulk of
inter-sample autocorrelation (slow rhythms in the delta / theta /
alpha range), and a higher-order predictor captures it with
correspondingly higher gain per added coefficient; the
highest-frequency detail subband is dominated by EMG and broadband
noise that adds little predictive structure beyond order 3, and
spending more coefficients there merely raises header overhead.
The orders were swept empirically and converged at this
schedule across the seven TUH corpora. An optional Adaptive mode
walks the Levinson--Durbin recursion up to a ceiling
of $P_{\max}=16$ and selects the AIC-optimal order per window
where the bench oracle indicates the fixed schedule is
sub-optimal (typically TUEV-style narrow-range recordings).

Bias cancellation following~\cite{sriraam2008} removes the DC offset
from the residual distribution prior to entropy coding. This
lightweight step empirically outperforms more complex neural
predictors on 250\,Hz 16-bit clinical EEG while adding negligible
computational cost---an important property for MCU deployment.
Three LPC modes are supported: Fixed (static coefficients),
Adaptive (per-window coefficient estimation), and Anytime (adaptive
with a per-subband deadline check for real-time operation); Anytime
is the default.

\subsection{Golomb-Rice Entropy Coding}
Golomb-Rice (GR) coding is theoretically optimal for
geometrically-distributed non-negative integers and near-optimal for
Laplacian-distributed residuals under appropriate parameter
selection~\cite{golomb1966,rice1979}. LPC residuals on EEG subbands
approximate a two-sided geometric (Laplacian-like) distribution,
making GR the natural entropy coder for this pipeline.

The Rice parameter $k$ is selected per channel per window from
residual statistics and stored in the fixed per-window header,
requiring no external dictionary. This design eliminates dictionary
overhead. rANS and arithmetic coding were evaluated and do not
recover their implementation overhead at the window sizes and
entropy levels characteristic of 250\,Hz 16-bit clinical EEG. This is consistent with the theoretical result that GR
achieves near-optimal coding efficiency for geometric sources with
lower per-symbol overhead than asymmetric numeral systems at short
block lengths. Windows are fully self-contained: no inter-window
state dependency exists, boundary effects are eliminated, and any
window can be decoded independently enabling random-access operation.

\subsection{LML Wire Format and Forensic Parity}
Each compressed window is packetized as a fixed 22-byte header
followed by a per-channel LPC metadata section and a variable-length
Golomb-Rice payload. The header is structured as: 4-byte magic
identifier (\texttt{LML1}); 2-byte channel count $n_{\text{ch}}$;
2-byte per-channel sample count $T$; 1-byte DWT level count;
1-byte flags; 4-byte LPC-metadata length; 4-byte payload length;
and 4-byte CRC-32 (standard 0x04C11DB7 polynomial, HD=6 at the
$\leq$\,4\,KiB window sizes used here per
Koopman~\cite{koopman2002crc}) computed over the LPC metadata and
payload,
guaranteeing bit-exact reconstruction per window. The LPC metadata
section that follows the header carries the per-channel, per-subband
predictor coefficients $a_i$, the chosen LPC order, and the Rice
parameter $k$ selected for that channel-subband pair. The
Golomb-Rice payload immediately follows. At the file level, the LML
container begins with the magic byte sequence and version field,
followed by channel metadata and a window offset table (LMLFOOT1
footer) enabling random-access decoding without a full-file scan. The wire format is integer-only and produces byte-identical
output across x86, ARM, and RISC-V implementations, verified by a
cross-platform consistency test suite.

Full-file forensic parity is maintained beyond signal content:
original EDF timestamps, annotations, and sidecar files are preserved
byte-exactly through the encode/decode round-trip. The only
permissible difference from the source directory is the OS-level
access time (\texttt{atime}), which is updated by the filesystem on
read. This zero-trust fidelity standard is motivated by clinical
deployment requirements: audit trails, medicolegal admissibility, and
FDA archival compliance depend on byte-identical reconstruction of
the complete recording, not merely the signal samples.

\subsection{Algebraic Structure: A Commutative Monoid}
\label{sec:monoid}
The LML wire format described above forms a \emph{commutative
monoid} under window concatenation. The identity element is the
empty sequence of windows. The binary operation is window
concatenation. The LMLFOOT1 footer makes the window set
self-describing: each window carries its own CRC-32, per-channel
LPC metadata, and timestamp in the LPC meta block. Associativity
follows from byte concatenation. Commutativity follows because
window order in the bitstream does not affect decoded signal
content---the footer index reassembles windows by timestamp at
decode time. Three deployment-relevant properties follow:
\begin{itemize}
\item \textbf{Embarrassingly parallel encoding.} Each window is
  encoded independently; the per-window bitstreams concatenate
  without inter-window synchronisation. The host-side encoder
  backend scales linearly with core count, and the firmware-side
  encoder produces a per-window unit that the basestation may
  ingest at its own pace.
\item \textbf{Append-without-re-encode.} Adding a new recording
  to an existing LMA archive requires concatenating the new
  window set and updating the footer index in place; prior content
  is never re-encoded. This is load-bearing for continuous
  acquisition where archives grow per-session over years.
\item \textbf{Order-independent decode and local fault containment.}
  Because each window self-validates via CRC-32 and self-locates via
  timestamp, the decoder may process windows out-of-order or in
  parallel. A corrupt window invalidates only itself, not the
  surrounding stream; the decoder skips it and continues. This
  pairs naturally with the failure-mode analysis discussed in
  Section~\ref{sec:discussion}.
\end{itemize}
The byte-equality conformance gate
(\texttt{lamquant-core/tests/byte\_equal\_backends.rs}) verifies
that this algebraic structure is preserved across x86, ARM, and
RISC-V backends: any optimisation that breaks bit-level identity
on a per-window basis is therefore a wire-format change requiring
a golden-table refresh, not a quiet performance tweak. The
algebra also makes the codec amenable to distributed encoding
across multiple hosts and to streaming acquisition without
end-of-stream signalling, which together cover the two
deployment patterns (offline archival, real-time wireless) the
codec is built to serve.

\subsection{Asymptotic Complexity}
\label{sec:complexity}
For a per-channel window of $N$ samples processed at $L$ DWT
levels with LPC order $P$, the per-window encoder cost decomposes
as:
\begin{itemize}
\item \textbf{DWT lifting}: $O(N)$ per level $\times$ $L$ levels =
  $O(NL)$, dominated by integer add and right-shift. With $L=3$
  fixed in the codec this collapses to $O(N)$.
\item \textbf{LPC residual}: $O(NP)$ per subband. Summed across
  four subbands with orders $\{3,3,6,8\}$, the per-window cost is
  $O(N \bar{P})$ with $\bar{P} = 5$ (mean of the fixed schedule).
  In Adaptive mode, the Levinson--Durbin recursion adds
  $O(P_{\max}^{2})$ per subband once per window, amortised over
  $N$ samples.
\item \textbf{Bias cancellation}: $O(N)$ per subband.
\item \textbf{Golomb-Rice encode}: $O(N)$ per subband, with a
  single \texttt{cpop} (Hazard3 Zbb~\cite{rp2350_datasheet}) per residual.
\item \textbf{CRC-32 + header pack}: $O(N + n_{\text{ch}})$ per
  window.
\end{itemize}
Aggregate: encoder cost is $O(N \bar{P})$ per window per channel,
i.e.\ linear in window size and proportional to mean LPC order.
For 32 channels at 250\,SPS, 10\,s windows ($N = 2500$,
$\bar{P} = 5$), the per-window operation count is on the order of
$32 \times 2500 \times 5 = 400{,}000$ MAC equivalents plus the
$O(N)$ linear passes --- well within real-time budget on a
150\,MHz MCU. Decoder cost is symmetric in the same big-$O$
class (inverse LPC, inverse lifting, GR decode) but runs on the
basestation host rather than the MCU.

\subsection{Decoder-Side Throughput}
\label{sec:decoder-throughput}
On the host x86 backend (Ryzen-class single core, AVX2 disabled to
keep parity with the firmware scalar path), the decoder sustains
$\approx$\,180\,MiB/s on the same window class as
Section~\ref{sec:results}.D --- $\approx$\,30\,\% faster than the
encoder, because the inverse-lifting stage is in-place
zero-allocation and the Golomb-Rice decode is a tight integer
bit-stream loop without coefficient computation. Decoder
throughput is not a deployment bottleneck for the
encoder-on-MCU + decoder-on-basestation split: even at maximum
archive ingest, a single basestation core decompresses faster
than the wireless link delivers compressed bytes.


% ---------------------------------------------------------------
% IV. EXPERIMENTAL RESULTS
% ---------------------------------------------------------------
\section{Experimental Results}
\label{sec:results}

\subsection{Datasets}
Experiments were conducted on two primary datasets.

\textbf{CHB-MIT Scalp EEG Database}~\cite{shoeb2009}: 23-channel
scalp EEG recordings from pediatric subjects with intractable
seizures, sampled at 256\,Hz, 16-bit resolution. The corpus
comprises 686 EDF files totalling 45,756,994,304 bytes (45.76\,GB);
LamQuant Lossless compresses it to 16,804,264,365 bytes
(16.80\,GB) under the adaptive LPC mode, yielding CR = 2.7229:1.
Used for direct comparison with Chen et~al.~\cite{chen2018vlsi}
under identical channel count and sampling rate parameters.

\textbf{Temple University EEG Corpus v2.0.2}~\cite{obeid2016,
tueg_aareadme_v202}: 70{,}831 EDF files from 14{,}987 patients
across 26{,}846 sessions (1.76\,TB, 27{,}074.4 hours,
250--1024\,Hz, 40+ channel configs). Released 2026-05-21;
v2.0.1 numbers retained in Appendix~A.
Bench timing: the full v2.0.1 benchmark was completed first, then
the v2.0.2 release became available mid-revision and the corpus
was updated via incremental rsync (only the v2.0.1 $\to$ v2.0.2
delta transferred, $\approx$\,32\,GB of $\approx$\,1.76\,TB
total); the per-montage CR bench was then re-run end-to-end on
the v2.0.2 tree to produce the headline numbers in
Table~\ref{tab:results}. Two files present in v2.0.0
were corrupted and rectified in v2.0.1; both were identified and
flagged correctly by the codec. The full verification run covered
88{,}147 encode/decode operations across 74{,}422 unique recordings
spanning the seven TUH EEG corpora (TUEG, TUSZ, TUAB, TUEP, TUEV,
TUAR, and TUSL), the CHB-MIT Scalp EEG Database, the
PhysioNet Siena Scalp EEG, Sleep-EDF, CAP Sleep, and EEG Motor
Movement/Imagery databases, and OpenNeuro ds004100. The per-operation
count exceeds the unique recording count because the labelled TUH
subsets (TUSZ, TUAR, TUAB, TUSL, TUEV, TUEP) re-list TUEG recordings
under their respective label montages, and the codec was run
independently against each sub-corpus to verify that label-aware
sidecars introduce no encode/decode regressions.

\subsection{Verification Methodology}
Bit-exact reconstruction was verified via four layers:
(1)~SHA-256 per file (matched on all 88{,}147 operations);
(2)~per-window CRC-32 during decompression;
(3)~pyedflib (C-backed reader) structural cross-check on 500
random TUEG files; (4)~MNE-Python signal-level cross-check on
$\approx$\,4{,}000 samples over multiple days. File counts and
directory structure were cross-referenced against the TUEG v2.0.2
AAREADME. Original EDF timestamps were confirmed preserved; the only
observed difference from source files was the OS-level \texttt{atime}
field. Zero failures were observed across 88{,}147 encode/decode
operations.

\subsection{Compression Results}
Table~\ref{tab:results} reports compression results on TUEG v2.0.2
by montage. The full corpus of 70{,}831 files
(1{,}756{,}355{,}590{,}458 bytes) compressed to
768{,}043{,}519{,}030 bytes, yielding a compression ratio of
2.287:1 and reducing the corpus from 1.76\,TB to 768\,GB
(measured by \texttt{tools/bench\_tueg\_subsets.py} in parallel
mode on a 20-core host, wall 49\,min). The v2.0.1 prior bench
(69{,}672 files, 1{,}760\,GB raw, 770\,GB compressed, CR 2.287:1)
is preserved in Appendix~A for traceability across the corpus
v2.0.1 $\rightarrow$ v2.0.2 transition. This constitutes the
first full-scale lossless compression benchmark on the full TUEG
v2.0.2 corpus.

\begin{table*}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{LamQuant Lossless: Compression Results on TUEG v2.0.2
by Montage}
\label{tab:results}
\centering
\begin{tabular}{lrrrr}
\toprule
Montage & Files & Original (GB) & Compressed (GB) & CR \\
\midrule
01\_tcp\_ar      & 51{,}361 & 1{,}373.05 & 603.11 & 2.277:1 \\
02\_tcp\_le      & 12{,}833 &   249.38   & 109.64 & 2.274:1 \\
03\_tcp\_ar\_a   &  6{,}621 &   133.70   &  55.19 & 2.422:1 \\
04\_tcp\_le\_a   &      16 &     0.2173 &   0.0937 & 2.318:1 \\
\midrule
\textbf{Full Corpus v2.0.2} & \textbf{70{,}831} & \textbf{1{,}756.36} & \textbf{768.04} & \textbf{2.287:1} \\
\bottomrule
\end{tabular}

\vspace{0.4em}
\noindent\parbox{0.95\textwidth}{\scriptsize $\dagger$~The
v2.0.1\,$\rightarrow$\,v2.0.2 transition added 1{,}159 EDF files
($+1.66$\,\%) while \emph{reducing} the on-disk raw byte total by
$\sim$\,3.7\,GB. This is consistent with the v2.0.2 AAREADME
changelog~\cite{tueg_aareadme_v202}: the release re-issued a
subset of v2.0.1 EDFs with corrected headers (typically tightening
the data-record byte count to drop trailing zero-padding) and
folded the prior TUSZ/TUEP edits back into TUEG. The aggregate
compression ratio is unchanged across the two releases (2.287:1
on both), confirming the transition is administrative rather than
distributional.}
\end{table*}

\subsection{Comparison with Prior Work}
Table~\ref{tab:comparison} compares LamQuant Lossless against
Chen et~al.~\cite{chen2018vlsi} and the gzip baseline. CHB-MIT
used Chen et~al.'s 23-channel 256\,Hz parameters. Three LPC
modes: fixed-schedule 2.6536:1, adaptive 2.7229:1, anytime
(adaptive + deadline gate) 2.7229:1. The adaptive mode is the
headline; the fixed-schedule number is closer in spirit to Chen
et~al.'s fixed-predictor architecture and still exceeds it by
12.9\,\%.

\begin{table*}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Comparison of Lossless EEG Compression Methods}
\label{tab:comparison}
\centering
\begin{tabular}{lllll}
\toprule
Method & Dataset & CR & PRD & MCU \\
\midrule
gzip $-9$~\cite{ziv1977,deutsch1996}
    & TUEG edf/000 & 1.6033:1 & 0\% & No \\
FLAC $-7$ (audio codec, fair per-channel)
    & TUEG edf/000 (456 files) & 2.2574:1 & 0\% & No \\
FLAC $-8$ (max LPC, fair per-channel)
    & TUEG edf/000 (456 files) & 2.2609:1 & 0\% & No \\
FLAC $-7$ (audio codec, fair per-channel)
    & CHB-MIT (686 files) & 2.6926:1 & 0\% & No \\
FLAC $-8$ (max LPC, fair per-channel)
    & CHB-MIT (686 files) & 2.6945:1 & 0\% & No \\
Chen et~al.~\cite{chen2018vlsi}
    & CHB-MIT & 2.35:1 & 0\% & Yes$^\dagger$ \\
\textbf{LamQuant (ours)}$^\ddagger$
    & CHB-MIT & \textbf{2.7229:1} & \textbf{0\%} & \textbf{Yes} \\
\textbf{LamQuant (ours)}
    & TUEG v2.0.2 & \textbf{2.287:1} & \textbf{0\%} & \textbf{Yes} \\
\bottomrule
\multicolumn{5}{l}{$^\dagger$Chen et~al.\ target a custom 0.18\,$\mu$m CMOS VLSI implementation. LamQuant targets commodity RP2350 RISC-V hardware.}\\
\multicolumn{5}{l}{$^\ddagger$LamQuant CHB-MIT is the adaptive LPC mode (best CR); fixed-schedule mode achieves 2.6536:1, anytime equals adaptive.}\\
\multicolumn{5}{p{0.95\textwidth}}{\scriptsize FLAC rows are benched per-channel (each EDF channel encoded as an independent 16-bit mono FLAC stream) because FLAC's native inter-channel decorrelation modes (mid/side, left/side, right/side) are designed for two-channel audio and degrade on uncorrelated 21-/23-channel EEG montages. The per-channel mono mode is the fairest comparator for a per-channel LPC codec; reviewers preferring the multi-channel fixed-prediction mode can substitute FLAC's \texttt{-{}-channel-assignment INDEPENDENT} flag without changing the result.}
\end{tabular}
\end{table*}

LamQuant Lossless achieves a 15.9\,\% improvement over Chen et~al.\
on their own benchmark dataset under matched evaluation conditions,
and establishes a 2.28:1 baseline on TUEG---a corpus an order of
magnitude larger than any previously used for lossless EEG
compression benchmarking. Figure~\ref{fig:cr-bar} summarises
compression ratios across all three methods and both evaluation
corpora.

\begin{figure}[!t]
\centering
\begin{tikzpicture}
\begin{axis}[
    xbar,
    width=0.95\columnwidth, height=4.6cm,
    bar width=8pt,
    enlarge y limits=0.18,
    xlabel={Compression Ratio (CHB-MIT, 686 files, 45.76\,GB)},
    xlabel style={font=\scriptsize},
    symbolic y coords={Chen,FLAC-7,FLAC-8,LamQuant},
    ytick=data,
    yticklabel style={font=\scriptsize},
    xticklabel style={font=\scriptsize},
    xmin=2.0, xmax=2.9,
    nodes near coords,
    every node near coord/.append style={font=\tiny},
    grid=major,
    major grid style={line width=0.2pt, draw=gray!30},
]
\addplot[fill=black!80, postaction={pattern=north east lines, pattern color=white}]
  coordinates {(2.35,Chen) (2.6926,FLAC-7) (2.6945,FLAC-8) (2.7229,LamQuant)};
\end{axis}
\end{tikzpicture}
\caption{Lossless compression on CHB-MIT
(\cite{shoeb2009}, 686 files, 45.76\,GB) — every comparator has a
real measurement on this corpus. LamQuant Lossless leads Chen
et~al.~\cite{chen2018vlsi} by 15.9\,\% and FLAC \texttt{-8} by
1.1\,\%. \textbf{The headline TUEG v2.0.2 result (2.287:1 over
70{,}831 files, 1.76\,TB raw) is the first lossless-EEG benchmark
at full-corpus scale and has no prior-art comparator;} see
Table~\ref{tab:results} and \S\ref{sec:scale}.}
\label{fig:cr-bar}
\end{figure}

\subsection{Throughput and Memory Footprint}
Table~\ref{tab:speed-mem} reports LamQuant Lossless throughput and
working-memory footprint along the two axes that matter for a
deployed wearable codec: encode/decode throughput (in samples or
MiB per second aggregated across all channels) and on-chip
working memory. Throughput is measured on host x86 with the
scalar firmware backend (the path that runs on RP2350
byte-for-byte) and SRAM footprint from per-stage buffer analysis
(\S\ref{sec:results}.D). Chen et~al.~\cite{chen2018vlsi} report
only an ASIC gate count and a clock-cycle figure derived from a
0.18\,$\mu$m CMOS implementation, which is not directly
comparable to a software MCU encode rate; we therefore omit them
from this table and confine the Chen et~al.\ CR comparison to
Table~\ref{tab:comparison}.

\begin{table*}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Encode Throughput and Working-Memory Footprint
(RP2350 row measured via RTL-level Verilator simulation of the
official Hazard3 core; see \S\ref{sec:results}.F.2)}
\label{tab:speed-mem}
\centering
\begin{tabular}{lrrl}
\toprule
Method & Throughput & Memory (KB) & Target \\
\midrule
LamQuant (host scalar, \emph{measured}) & 137\,MiB/s & 278 & Ryzen-class x86 \\
LamQuant (host scalar decoder, \emph{measured}) & 180\,MiB/s & 278 & Ryzen-class x86 \\
LamQuant (RP2350 encoder, \emph{measured})$^\ddagger$ & 0.627\,Msa/s & 278 & Hazard3 150\,MHz \\
\bottomrule
\end{tabular}

\vspace{0.4em}
\noindent\parbox{0.95\textwidth}{\scriptsize $^\ddagger$ \textbf{Measured via
RTL-level Verilator simulation} of the official Hazard3 RTL
(\texttt{Wren6991/Hazard3} v1.1, the same Verilog that taped out
as the RP2350 RISC-V cores) using the bench harness at
\texttt{tools/hazard3\_bench/} (boots, builds a 21\,ch\,$\times$\,2500\,sample
deterministic xorshift window, runs \texttt{PipelineScheduler::
encode\_window} eight times under an \texttt{mcycle}+\texttt{minstret}
bracket). Measured: 12{,}548{,}067 cycles/window
$\Rightarrow$ 83.66\,ms/window\,@\,150\,MHz $\Rightarrow$
0.627\,Msa/s = 119$\times$ real-time for a 21\,ch\,@\,250\,Hz
input. CPI = 1.071 (achieved IPC 0.93, within 7\,\% of the IPC
1.0 ceiling for Hazard3's single-issue 3-stage pipeline).
On-silicon Pico 2 board confirmation pending.}
\end{table*}

\subsection{Embedded Deployment on RP2350}
\noindent\fbox{\parbox{\columnwidth}{\footnotesize
\textbf{Status — RP2350 deployment characterised via RTL
simulation.} The encode-latency number reported below is
\emph{measured} by running the encoder ELF on the official
\texttt{Wren6991/Hazard3} v1.1 RTL under
Verilator~\cite{verilator} (the same Verilog that taped out as
the RP2350 RISC-V cores), with the Hazard3 \texttt{mcycle} and
\texttt{minstret} CSRs read across the encode loop. Wire-format
parity is independently enforced by the
\texttt{byte\_equal\_backends} conformance gate; the
on-physical-silicon Pico 2 run is the only remaining step and
will be reported as an erratum or in the next revision.
}}

\vspace{0.4em}

The LML encoder targets the RP2350 microcontroller (dual Hazard3
RISC-V cores, 150\,MHz, 64\,KB SRAM banks)~\cite{rp2350_datasheet}.
The Hazard3 Zbb bit-manipulation extension provides hardware \texttt{cpop}
(population count) in a single cycle, accelerating the Golomb-Rice
inner loop relative to software lookup-table implementations on ARM
Cortex-M0/M3 (8--16 cycles/byte). The encoder-only design runs on
the MCU; the decoder executes on the base station, keeping the MCU
firmware footprint minimal. The integer-only wire format eliminates
floating-point hardware requirements entirely.

\subsubsection{Encode Latency}
A 21-channel 2500-sample window (52{,}500 Q31 samples in the
encoder hot path, 10\,s wall time at 250\,SPS per channel,
102.5\,KB raw i16 ADC payload before fixed-point conversion)
encodes in \textbf{12{,}548{,}067 cycles} on the Hazard3 RTL, measured by
the \texttt{tools/hazard3\_bench/bench\_encode} harness running
under Verilator against the official \texttt{Wren6991/Hazard3}
v1.1 RTL (same Verilog that taped out as RP2350). At the RP2350
core clock of 150\,MHz this is \textbf{83.66\,ms per window}, or
\textbf{0.627\,Msa/s} aggregate throughput. The 21\,ch\,@\,250\,Hz
input stream needs 10 seconds of wall time per 52{,}500-sample
window, so the encoder finishes \textbf{119{$\times$} ahead of
real-time}, leaving 9.92\,seconds of every 10-second window free
for SNN inference, BLE transmission, ADC streaming, and power
management. The achieved CPI is 1.071 (IPC 0.93)---within 7\,\% of the
theoretical maximum (IPC 1.0) for Hazard3's single-issue
3-stage in-order scalar pipeline---confirming that the codec
is predictor-bound rather than fetch- or memory-bound, and
that there is essentially no headroom to be gained from
microarchitectural pipelining tricks at this design point.

\subsubsection{Encoder SRAM Footprint}
The encoder working set on a 21-channel window is dominated by the
LPC residual buffer (21\,$\times$\,2500\,$\times$\,4\,B
= 205\,KB i32 residuals), the bias-cancellation context
(21\,$\times$\,256\,$\times$\,4\,B = 21\,KB), and a variable-size
Golomb-Rice output buffer (bounded by residual size, typically
20--50\,KB). The lifting DWT operates in-place on the input buffer
and adds no scratch beyond the working window. Total encoder
working memory is approximately 278\,KB (rounded down from
278.5\,KB; the firmware-level optional biquad high-pass
preprocessing adds a further 1.5\,KB of delay-line state but is
not part of the lossless codec proper). All numbers fit comfortably
in the RP2350's 520\,KB on-chip SRAM with headroom for safety state,
USB / BLE transport buffers, and the basestation-bound activity-map
output of the on-MCU neural front-end.

\subsubsection{Encoder-on-MCU + Decoder-on-Basestation Split}
The codec is structured so that encoding runs on the MCU (battery-
operated wearable) while decoding runs on the basestation
(line-powered host). This asymmetry reduces MCU flash footprint and
power consumption: the encoder does not carry the inverse lifting,
inverse LPC, or Golomb-Rice decode paths, freeing flash for AFE
drivers, safety state machines, and the on-MCU SNN seizure
detector. The wire format's byte-equality across architectures
(verified by a dedicated conformance gate) guarantees that the
basestation decoder reconstructs the original signal bit-for-bit
regardless of which host architecture the encoder ran on.


% ---------------------------------------------------------------
% V. DISCUSSION
% ---------------------------------------------------------------
\section{Discussion}
\label{sec:discussion}

The 2.287:1 result on TUEG exceeds both Shannon ceilings from
\S\ref{sec:background} ($H_0 \approx 10.775$\,bit/s,
$H_0(\Delta X) \approx 9.337$). This implies an effective residual
entropy of $\leq 7.0$\,bits/sample after the LamQuant pipeline,
i.e. the four-band lifting wavelet plus per-subband adaptive LPC
of order up to~8 captures considerably more inter-sample structure
than a first-order predictor. Artifact-heavy recordings and diverse
montage configurations suppress CR relative to curated laboratory
datasets; this is the correct behavior of an honest benchmark, not
a limitation of the codec. The result on TUEG should be interpreted
as a conservative lower bound on achievable CR for clean ambulatory
recordings.

Clinical deployment of a lossless EEG codec imposes requirements
beyond signal content fidelity. Medicolegal admissibility, FDA
archival compliance, and institutional audit trails require that
timestamps, annotations, sidecar files, and directory structure are
preserved byte-exactly across the encode/decode round-trip. LamQuant
Lossless adopts a zero-trust verification policy: SHA-256 per file,
CRC-32 per window, and independent cross-validation with multiple EDF
reader implementations. This standard exceeds the verification
requirements of prior lossless EEG compression work, which has
typically reported only signal-level PRD without addressing metadata
and annotation fidelity.

\subsection{Failure-Mode Analysis on Corrupted Inputs}
\label{sec:failure-mode}
The verification run encountered two file populations the
encoder declined to process. Two TUEG v2.0.0 files had record
counts short of their declared length (rectified by the corpus
maintainers in v2.0.1; both replacements processed cleanly). Second, the CAP Sleep corpus
(PhysioNet) contains two EDFs (of 108 total --- approximately
1.9\%) with malformed headers: \texttt{brux1.edf} has a
declared file size that does not match its data record count,
and \texttt{ins1.edf} has an out-of-range Digital Minimum
field. Both files are upstream-public release defects rather than
local corruption. LamQuant Lossless's encoder declines to silently
truncate, pad, or coerce malformed inputs; the per-window CRC-32
plus the per-file SHA-256 chain surface the discrepancy at archive
time, and the offending stem is flagged in the build log rather
than silently corrupted. This behavior is the correct one for a
clinical archival codec: the codec must never \emph{invent}
samples to make a malformed input fit a clean output frame.

\subsection{When LamQuant Compresses Poorly}
\label{sec:poor-compression}
Distinct from input-corruption failures, the codec also has
\emph{recording-type} failure modes worth documenting. The lowest
per-file CR observed in the full TUEG\,edf/000 benchmark (456
files) was 1.05:1 on \texttt{aaaaaacz/s007\_2015/01\_tcp\_ar/
aaaaaacz\_s007\_t006.edf} (4.5\,MB raw); the same subject's
session 7 contributes three of the five worst-CR files in the
edf/000 cohort (CR = 1.05, 1.30, 1.69 on tokens
t006, t000, t005 respectively). Visual inspection of these
recordings shows a near-uniform high-amplitude broadband
disturbance --- consistent with a disconnected reference
electrode or persistent EMG / mains contamination saturating the
16-bit range across all channels. The lifting DWT cannot
decorrelate noise this broad-spectrum, and the LPC residuals
inherit the same near-white statistics; Golomb-Rice on a
near-uniform residual yields very little gain over the raw
representation. The codec is correct on these files (bit-exact
roundtrip via the conformance gate); it is simply at its
information-theoretic floor.

These cases inform the realistic CR floor that downstream system
designers should assume: even a well-behaved lossless codec will
sit close to 1:1 on individual recordings dominated by
broadband artifact. The corpus-average CR claimed in
Table~\ref{tab:results} (2.287:1 over 70{,}831 files) is the
relevant number for archival sizing; designers should not assume
the average applies to every individual recording.

\subsection{Why Benchmark Scale Matters}
\label{sec:scale}
Per-file CR distributions on the 13 benched corpora highlight a
methodological point: \emph{compression ratios reported on small
datasets systematically overstate the achievable CR on larger,
more diverse corpora}. On TUEG edf/000 (456 files,
6.0\,GB) the per-file CR distribution
(\texttt{tools/bench\_per\_file\_cr.py}) ranges from 1.05 to
6.51 (median 2.29, $\sigma = 0.66$); a single recording achieves
6.51:1 while the aggregate over the full subset is 2.299:1, and the
aggregate over the full 1.76\,TB TUEG corpus is 2.287:1 ---
essentially unchanged from the 456-file subset, because the
per-file outliers are averaged out. CHB-MIT (45.76\,GB, 686 files)
shows the same pattern: the per-file CR ranges 1.72 to 5.14
(median 2.82), and the corpus aggregate is 2.723:1 --- a respectable
number, but 19\% lower than the TUEG single-file maximum of 6.51.
The OpenNeuro intracranial dataset achieves 3.573:1 aggregate
because its 319 recordings come from a single hospital with a
common acquisition setup; if it were ten times larger and drawn
from multiple sites with diverse hardware, the aggregate would
almost certainly drop. The corollary for the reader of any
lossless EEG paper is concrete: \emph{benchmark CR scales inversely
with corpus diversity}, and headline numbers from sub-100-file
benches should be discounted accordingly. LamQuant Lossless's
1.76\,TB TUEG result, taken at the scale of an entire clinical EMU
corpus, is therefore the load-bearing benchmark; the CHB-MIT
2.723:1 enables direct comparison with prior work but is intrinsically
a higher number than the full-scale clinical mixture justifies.

\begin{figure*}[!t]
\centering
\begin{tikzpicture}
\begin{axis}[
    boxplot/draw direction=y,
    width=0.95\textwidth,
    height=7cm,
    ymin=0.8, ymax=7.5,
    ylabel={Per-file CR (LamQuant Lossless)},
    ylabel style={font=\small},
    xtick={1,2,3,4,5,6},
    xticklabels={
        {TUEG sample},
        {CHB-MIT},
        {TUEV},
        {Sleep-EDF},
        {EEG MMI},
        {ds004100 (iEEG)},
    },
    xticklabel style={font=\small},
    yticklabel style={font=\small},
    ytick={1,2,3,4,5,6,7},
    minor y tick num=1,
    grid=both,
    major grid style={line width=0.25pt, draw=gray!40},
    minor grid style={line width=0.1pt, draw=gray!20},
    cycle list={{black}},
    boxplot/box extend=0.5,
    boxplot/whisker extend=0.35,
]
% Whiskers use p5 / p95 (statistically robust); min/max marked as
% outliers via custom coordinates below. Each addplot below also
% pins the aggregate CR as a thick dashed line via "median" trick
% — but to keep clarity we just annotate the aggregate above each box.
\addplot+[boxplot prepared={
    lower whisker=1.870, lower quartile=2.113, median=2.274,
    upper quartile=2.493, upper whisker=2.891,
}, fill=gray!30, draw=black, line width=0.5pt] coordinates {};
\addplot+[boxplot prepared={
    lower whisker=2.156, lower quartile=2.487, median=2.820,
    upper quartile=3.244, upper whisker=3.918,
}, fill=gray!30, draw=black, line width=0.5pt] coordinates {};
\addplot+[boxplot prepared={
    lower whisker=2.540, lower quartile=3.188, median=3.500,
    upper quartile=3.798, upper whisker=4.282,
}, fill=gray!30, draw=black, line width=0.5pt] coordinates {};
\addplot+[boxplot prepared={
    lower whisker=1.877, lower quartile=1.961, median=2.032,
    upper quartile=2.209, upper whisker=3.046,
}, fill=gray!30, draw=black, line width=0.5pt] coordinates {};
\addplot+[boxplot prepared={
    lower whisker=2.211, lower quartile=2.391, median=2.559,
    upper quartile=2.904, upper whisker=3.627,
}, fill=gray!30, draw=black, line width=0.5pt] coordinates {};
\addplot+[boxplot prepared={
    lower whisker=2.648, lower quartile=3.177, median=3.577,
    upper quartile=3.945, upper whisker=4.379,
}, fill=black!75, draw=black, line width=0.5pt] coordinates {};
% Min/max outlier markers (filled circles)
\addplot[only marks, mark=*, mark size=1.2pt, black] coordinates {
    (1, 1.127) (1, 6.915)
    (2, 1.723) (2, 5.139)
    (3, 1.190) (3, 5.375)
    (4, 1.360) (4, 3.434)
    (5, 1.923) (5, 4.255)
    (6, 2.312) (6, 5.120)
};
% Aggregate CR markers (diamond), one per corpus
\addplot[only marks, mark=diamond*, mark size=3pt,
         fill=white, draw=black, line width=0.8pt] coordinates {
    (1, 2.259) (2, 2.723) (3, 3.409)
    (4, 2.087) (5, 2.638) (6, 3.573)
};
\end{axis}
\end{tikzpicture}
\caption{Per-file CR distribution across six representative
corpora (3{,}964 files total: \textbf{200-file random TUEG
sample}, 686 CHB-MIT, 1{,}036 TUEV, 197 Sleep-EDF, 1{,}526 EEG
MMI DB, 319 OpenNeuro ds004100 iEEG). Box = IQR (p25--p75) with
median bar; whiskers = p5--p95 (statistically robust); filled
circles = per-file extremes (min, max); open diamonds = aggregate
CR (sum-raw / sum-compressed) per corpus. \emph{The TUEG diamond
shows the 200-file sample aggregate (2.259); the full-corpus
TUEG v2.0.2 aggregate over all 70{,}831 files is 2.287:1, reported
in Table~\ref{tab:results} as the headline TUEG number. The
200-file sample reaches a per-file maximum of 6.92:1 on a
low-noise resting recording; the full edf/000 subset (456 files)
has a maximum of 6.51:1, also reflected in Section~\ref{sec:scale}.}
iEEG (filled black box) sits highest by aggregate
(3.573:1) and by p25 (3.18). TUEV's elevated median (3.5:1)
reflects its narrower dynamic range (alphabet 15{,}416 of 65{,}536
possible 16-bit codes; see Table~\ref{tab:per-corpus-entropy}).}
\label{fig:per-file-cr}
\end{figure*}

\subsection{Scope Exclusions}
\textbf{(i) Sampling rate \& bit depth}: benchmarking used
corpus-native rates (250--1024\,Hz TUEG; up to 16\,kSPS on some
PhysioNet datasets) and 16-bit resolution. Performance at
higher sampling rates (e.g., 2\,kSPS for high-frequency oscillation
detection) and 24-bit resolution at scale has not yet been
characterised; the codec is parameter-agnostic at the algorithm
level, but the published CR figures are not portable to those
regimes without re-measurement.
\textbf{(ii) Modality coverage}: scalp EEG is the headline target
(twelve of the thirteen benched corpora). One intracranial corpus
(OpenNeuro ds004100) is included to characterise the codec on
iEEG; the higher CR (3.573:1) reflects the tighter spectral
content of intracranial recordings vs scalp, and per-window LPC
order selection adapts accordingly. High-density ECoG and
multi-thousand-channel sEEG arrays remain to be characterised.
\textbf{(iii) Patient-level metadata}: LamQuant Lossless preserves
per-recording metadata (annotations, sidecar files, EDF headers)
byte-exactly. Multi-recording-per-patient or longitudinal
aggregation across patient sessions is out of scope for this work;
that layer belongs to the dataset / corpus management tooling, not
the per-recording codec.

\subsection{Why LPC, Not Neural Predictors}
\label{sec:why-lpc}
Neural-network predictors have been proposed as alternatives to
linear prediction for EEG compression~\cite{sriraam2012}, offering
the potential to capture higher-order inter-sample dependencies
that a fixed-order linear predictor cannot. The choice in LamQuant
Lossless to use per-subband adaptive LPC (a fixed schedule of
orders $\{3, 3, 6, 8\}$ with an optional adaptive mode of maximum
order~16) instead is grounded in three constraints that are load-
bearing for a clinically deployable codec.

\emph{Domain generalization.} A neural predictor trained on
TUEG may underperform on CHB-MIT, Sleep-EDF, Siena, or future
corpora with different hardware, montages, electrode impedance,
or patient populations. Per-subband adaptive LPC, by contrast, fits
its coefficients from the local signal within each window via the
Levinson-Durbin recursion. There is no training set and therefore
no domain shift; the predictor is universal across corpora by
construction. The 88{,}147-operation verification run reported in
Section~\ref{sec:results} spans thirteen heterogeneous EEG datasets
with zero per-file CR pathology attributable to corpus mismatch.

\emph{Microcontroller deployability.} The encoder must execute in
real time on a 150~MHz Hazard3 RISC-V core within $\leq$278\,KB
of SRAM working set (Section~\ref{sec:results}). Linear prediction
with the fixed-schedule $\{3, 3, 6, 8\}$ ladder fits this envelope
trivially: the per-window coefficient state is 20~scalars and the
per-sample inner loop is at most eight multiply-accumulate
operations. Neural predictors that would meaningfully improve over
LPC require quantized weight storage in flash and inference compute
that scales with hidden dimension and context length, neither of
which is bounded by training-time accuracy goals alone. We have
prototyped quantized predictor architectures compatible with the
RP2350 deployment envelope (1.58-bit ternary weight quantization
in the manner of Ma et~al.~\cite{ma2024bitnet}) and have observed
marginal CR improvement (sub-1\%) relative to LPC at substantial
flash and latency cost. The improvement does not justify the deployability
penalty for a codec whose primary requirement is universality, not
last-percent CR.

\emph{Audit and certification.} The fixed-coefficient form of the
LPC inner loop is amenable to byte-equality conformance gates
across x86, ARM, and RISC-V backends (the
\texttt{byte\_equal\_backends} test in
\texttt{lamquant-core/tests/}). Replacing this loop with a neural
predictor introduces non-determinism that complicates regulatory
audit, because the trained weights are themselves a deployment
artifact subject to change.

Closing the residual entropy gap between LPC and the lower bound
implied by $H_0(\Delta X) \approx 9.337$\,bits/sample (and the even
lower per-corpus values reported in Section~\ref{sec:background}
and Appendix~A) is therefore left as future work, contingent on
predictor architectures that satisfy the three constraints above
simultaneously.


% ---------------------------------------------------------------
% VI. CONCLUSION
% ---------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}

LamQuant Lossless achieves a compression ratio of 2.72:1 on the
CHB-MIT Scalp EEG Database---a 15.9\,\% improvement over the prior
state of the art under identical evaluation conditions---and 2.28:1
on the Temple University EEG Corpus v2.0.2 — the
first full-scale lossless compression benchmark on the largest
open clinical EEG dataset in existence. Bit-exact reconstruction was verified across
88{,}147 encode/decode operations on 74{,}422 unique recordings
spanning thirteen clinical and research EEG datasets with zero failures. The codec is
integer-only, dictionary-free, and deployable in real time on
commodity RISC-V microcontroller hardware without floating-point
support. It is released open-source as the archival compression layer
of the OH-EEG wireless EEG platform.


% ---------------------------------------------------------------
% ACKNOWLEDGMENTS / STATEMENTS
% ---------------------------------------------------------------
\section*{Acknowledgments}
The author thanks the Neural Engineering Data Consortium at Temple
University for releasing the TUH EEG Corpus, and the PhysioNet
project for hosting the CHB-MIT Scalp EEG Database, Siena Scalp
EEG, Sleep-EDF, CAP Sleep, and EEG Motor Movement/Imagery
databases used as comparators in this work.

\section*{Funding}
This work was self-funded by OpenHuman Technologies LLC, Florida,
USA. No external grants or external sponsorship contributed to
the design, execution, or publication of this research.

\section*{Conflict of Interest}
The author is the founder of OpenHuman Technologies LLC and has
filed a (provisional) patent application related to the codec
described in this paper (Patent Pending US \#64/032,641). The
source code is released open-source under the GNU Affero General Public
License version 3 or later (AGPLv3-or-later)~\cite{agpl3} at
\url{https://github.com/Quitetall/LamQuant-Lossless} and permanently
archived at Zenodo (DOI:
\href{https://doi.org/10.5281/zenodo.20436114}{10.5281/zenodo.20436114}).
The AGPLv3 license
includes an automatic patent grant under \S 11, conveying an
express patent license from every contributor to every downstream
recipient covering any claims that read on the released code.
This is materially stronger than a unilateral commitment not to
enforce: every user of the AGPLv3-licensed implementation already
holds an irrevocable patent license for the act of using,
modifying, and redistributing that implementation, by operation
of the license terms.

\section*{Ethics Statement}
This study uses only publicly released, fully de-identified EEG
corpora that were approved for open distribution by their
originating institutions. No human-subject experimentation,
clinical intervention, or new data collection was performed by
the author. No IRB approval was required for the benchmarking
described herein.

\section*{Reproducibility and Data Availability}

\subsection*{Reproducibility Package}
Every numerical claim in this paper is reproducible from the
published source tree:

\begin{itemize}
\item \textbf{Codec source}: \texttt{lamquant-core/} (Rust, no\_std)
  and \texttt{lamquant-firmware/} (RP2350 bare-metal bin).
\item \textbf{Bench scripts}: \texttt{tools/bench\_chbmit.py},
  \texttt{tools/bench\_tueg\_subsets.py},
  \texttt{tools/bench\_per\_file\_cr.py},
  \texttt{tools/bench\_shannon\_entropy.py},
  \texttt{tools/bench\_gzip\_baseline.sh},
  \texttt{tools/bench\_edf\_reader\_parity.py}.
\item \textbf{Verification gate}: \texttt{tools/verify\_paper\_claims.py}
  cross-checks every numerical claim against the
  \texttt{evidence/*.json} files and returns \texttt{60 PASS / 0
  FAIL} at submission. The PASS-tally is a precondition for
  submission.
\item \textbf{Byte-equality conformance gate}:
  \texttt{lamquant-core/tests/byte\_equal\_backends.rs} verifies
  that the host scalar, host vectorised, and firmware backends
  produce byte-identical \texttt{.lml} output for every golden
  vector — load-bearing for the per-corpus deterministic numbers.
\item \textbf{RP2350 silicon bench harness}:
  \texttt{tools/hazard3\_bench/} ships a ready-to-flash
  \texttt{bench\_encode} ELF that reads \texttt{mcycle} /
  \texttt{minstret} CSRs across an \texttt{encode\_window} loop;
  the on-silicon runbook (Verilator-on-RTL, CXXRTL-on-RTL, and
  real-board paths) lives at \texttt{tools/bench\_rp2350\_silicon.md}.
\end{itemize}

\subsection*{Data Availability}
All EEG corpora used in this study are publicly available without
restriction or registration via the following sources:

\begin{itemize}
\item TUH EEG Corpus (v2.0.2): \\
  \url{https://isip.piconepress.com/projects/tuh_eeg/}
\item CHB-MIT Scalp EEG Database: \\
  \url{https://physionet.org/content/chbmit/1.0.0/}
\item PhysioNet Siena Scalp EEG, Sleep-EDF, CAP Sleep, EEG MMI DB:\\
  \url{https://physionet.org/}
\item OpenNeuro ds004100 (iEEG, HUP):\\
  \url{https://openneuro.org/datasets/ds004100}
\end{itemize}

The LamQuant Lossless codec source code, all benchmark scripts,
per-corpus JSON evidence files, and reproducibility documentation
are released open-source at
\url{https://github.com/Quitetall/LamQuant-Lossless} under the
\texttt{docs/paper/} and \texttt{tools/} directories. All numbers
reported in this paper can be reproduced from these scripts on a
local mirror of the listed corpora.

\section*{Author Contributions}
Per the Contributor Roles Taxonomy (CRediT):
\textbf{Brian Lam}: Conceptualization, Methodology, Software,
Validation, Formal analysis, Investigation, Resources, Data
Curation, Writing -- Original Draft, Writing -- Review \& Editing,
Visualization, Supervision, Project administration, Funding
acquisition.


% ---------------------------------------------------------------
% REFERENCES (placed before the appendix so the bibliography
% flows continuously out of the backmatter sections; the appendix
% tables + biography close the paper without orphaning the final
% reference under unrelated tables on a later page).
% ---------------------------------------------------------------
\bibliographystyle{IEEEtran}
\bibliography{references}

% ---------------------------------------------------------------
% APPENDIX A — Supplementary per-corpus breakdown
% ---------------------------------------------------------------
\appendices

\section{Per-corpus Compression Breakdown}
\label{app:per-corpus}

Table~\ref{tab:per-corpus} breaks the headline TUEG number into
per-corpus rows for every TUH EEG dataset plus the CHB-MIT
PhysioNet release. All numbers use the default configuration
(Anytime LPC, three-level Le~Gall 5/3 lifting DWT, integer wire
format, per-window CRC-32). Smaller corpora compress better
because their content (artifact, brief seizures, controlled
task experiments) is narrower than TUEG's open-ended EMU stream.

\begin{table*}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Per-Corpus Compression on the 13 Benched Datasets}
\label{tab:per-corpus}
\centering
\begin{tabular}{lrrrl}
\toprule
Corpus & Files & Input (GB) & Compressed (GB) & CR \\
\midrule
TUSL v2.0.1 &   112 &    1.63 &   0.70 & 2.311:1 \\
TUAR v3.0.1 &   489 &    9.28 &   4.15 & 2.235:1 \\
TUEV v2.0.1 & 1{,}036 &  16.05 &   4.71 & \textbf{3.409:1} \\
TUAB v3.0.1 & 4{,}181 &  87.88 &  38.55 & 2.280:1 \\
TUEP v3.1.0 & 2{,}255 &  31.94 &  14.03 & 2.276:1 \\
TUSZ v2.0.6 & 8{,}141 &  84.67 &  35.79 & 2.366:1 \\
\textbf{TUEG v2.0.2}    & \textbf{70{,}831} & \textbf{1{,}756.36} & \textbf{768.04} & \textbf{2.287:1} \\
TUEG v2.0.1$^\ast$ & 69{,}672 & 1{,}760.07 & 769.68 & 2.287:1 \\
\midrule
\multicolumn{5}{l}{\textit{External (scalp EEG):}} \\
CHB-MIT             &    686  & 45.76 & 16.80 & 2.723:1 \\
Siena Scalp EEG     &     41  & 21.80 &  8.62 & 2.528:1 \\
Sleep-EDF           &    197  &  8.71 &  4.18 & 2.087:1 \\
CAP Sleep$^\dagger$ &    106  & 42.81 & 18.05 & 2.371:1 \\
EEG MMI DB          &  1{,}526 &  3.61 &  1.37 & 2.638:1 \\
\midrule
\multicolumn{5}{l}{\textit{External (intracranial EEG):}} \\
OpenNeuro ds004100$^\ddagger$ & 319 & 14.19 & 3.97 & \textbf{3.573:1} \\
\bottomrule
\end{tabular}

\vspace{0.4em}
\noindent\footnotesize
$^\ast$ TUEG v2.0.1 row retained for traceability across the
v2.0.1 $\rightarrow$ v2.0.2 corpus transition; v2.0.2 is the
headline measurement.
\hfill\break
$^\dagger$ CAP Sleep has 2 EDF files (out of 108) with malformed
headers that the LamQuant encoder declines; see
Section~\ref{sec:failure-mode}.
\hfill\break
$^\ddagger$ Only intracranial EEG corpus in the bench set;
substantially higher CR reflects the tighter spectral content
of iEEG vs scalp recordings.

\end{table*}

\subsection*{Per-Corpus Shannon Entropy and Compression Ceilings}

Table~\ref{tab:per-corpus-entropy} reports empirical Shannon
entropies for each corpus, computed by
\texttt{tools/bench\_shannon\_entropy.py} on a random sample (seed
\texttt{0x4C414D51}) of up to 200 EDFs per corpus, contributing
every sample of every channel to a single global histogram per
corpus. $H_0(X)$ is the zero-order entropy of the raw 16-bit
digital sample stream; $H_0(\Delta X)$ is the zero-order entropy
of the first-order difference (LPC-1 residual). The corresponding
CR ceilings are $16/H$.

\begin{table*}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Per-Corpus Shannon Entropy on the 13 Benched Corpora
($n \approx 56.3 \times 10^9$ samples)}
\label{tab:per-corpus-entropy}
\centering
\begin{tabular}{lrrrrr}
\toprule
Corpus & Files & $H_0(X)$ & $H_0(\Delta X)$ & CR$_{\text{raw}}$ ceil & CR$_{\Delta}$ ceil \\
\midrule
TUAR v3.0.1 & 489 & 9.862  & 8.327 & 1.622:1 & 1.922:1 \\
TUEV v2.0.1 & 200 & 7.063  & 5.620 & 2.265:1 & 2.847:1 \\
TUSL v2.0.1 & 112 & 10.402 & 8.327 & 1.538:1 & 1.922:1 \\
TUSZ v2.0.6 & 200 & 10.412 & 8.865 & 1.537:1 & 1.805:1 \\
\textbf{TUEG v2.0.2} & \textbf{200} & \textbf{10.795} & \textbf{9.018} & \textbf{1.482:1} & \textbf{1.774:1} \\
TUEG v2.0.1$^\ast$ & 200 & 10.775 & 9.337 & 1.485:1 & 1.714:1 \\
TUAB v3.0.1 & 200 & 9.499  & 7.885 & 1.684:1 & 2.029:1 \\
TUEP v3.1.0 & 200 & 10.733 & 9.564 & 1.491:1 & 1.673:1 \\
\midrule
\multicolumn{6}{l}{\textit{External:}} \\
CHB-MIT (scalp)     & 200 & 8.387  & 6.724 & 1.908:1 & 2.379:1 \\
Siena Scalp EEG     &  41 & 10.509 & 8.312 & 1.522:1 & 1.925:1 \\
Sleep-EDF           & 196 & 9.343  & 8.022 & 1.713:1 & 1.994:1 \\
CAP Sleep           &  96 & 10.671 & 7.414 & 1.499:1 & 2.158:1 \\
EEG MMI DB          & 200 & 7.897  & 6.648 & 2.026:1 & 2.407:1 \\
OpenNeuro ds004100 (iEEG) & 128 & 9.755 & 5.378 & 1.640:1 & 2.975:1 \\
\midrule
\textbf{Weighted aggregate} & --- & \textbf{10.066} & \textbf{7.659} &
\textbf{1.590:1} & \textbf{2.089:1} \\
\bottomrule
\end{tabular}

\vspace{0.4em}
\noindent\footnotesize
$^\ast$ TUEG v2.0.1 row retained for traceability of the prior
bench across the v2.0.1 $\rightarrow$ v2.0.2 corpus transition;
the aggregate uses v2.0.1 only to avoid double-counting TUEG.
The weighted aggregate is computed over all benched samples
across the 13-corpus mixture ($\approx 56.3 \times 10^9$
samples total). The supplementary
\texttt{shannon\_entropy\_full\_summary.json} aggregate
($H_0(X) = 9.244$, $H_0(\Delta X) = 7.656$ over
$16.8 \times 10^9$ samples) is over the 200-file-per-corpus
subsample used by \texttt{tools/verify\_paper\_claims.py} and
intentionally weights smaller corpora more heavily for
reproducibility on a downloader's local mirror.
\end{table*}

Two observations follow. First, $H_0(\Delta X)$ is consistently
1--3\,bits/sample lower than $H_0(X)$, confirming substantial
sample-to-sample correlation that any first-order predictor can
exploit. Second, even the $\text{CR}_{\Delta}$ ceiling (the best a
zero-order entropy coder can do over LPC-1 residuals) sits below
LamQuant's measured 2.287:1 on TUEG and 2.7229:1 on CHB-MIT,
which means LamQuant's four-band lifting DWT plus per-subband
adaptive LPC of order up to~8 captures higher-order structure
beyond what first-order prediction reaches. The unusually low
entropies on TUEV ($H_0 = 7.06$, alphabet 15{,}416 of 65{,}536
possible 16-bit codes) reflect the narrower dynamic range of
event-classification recordings relative to open-ended
EMU data. TUAB and TUEP are pending an integrity-driven
re-rsync from the TUH server; their entropies will be added in
a revision.

The codec's per-window structure makes any subset of the corpus
trivially re-bench-able. Reproducer scripts are released alongside
the codec source at
\url{https://github.com/Quitetall/LamQuant-Lossless} (\texttt{tools/
bench\_chbmit.py}, \texttt{tools/bench\_gzip\_baseline.sh},
\texttt{tools/bench\_tueg\_subsets.py},
\texttt{tools/bench\_shannon\_entropy.py}).


% ---------------------------------------------------------------
% BIOGRAPHY
% ---------------------------------------------------------------
\begin{IEEEbiographynophoto}{Brian Lam}
Brian Lam is the founder of OpenHuman Technologies LLC, Florida, USA,
where he is developing an open-source modular wireless EEG acquisition
platform targeting clinical seizure monitoring and brain-computer
interface applications. His research interests include biosignal
compression, embedded machine learning, and medical device engineering.
\end{IEEEbiographynophoto}

\end{document}
